---
title: "leave_one_out() and orchard_leave1out() examples"
author: "fd"
date: "2025-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning=FALSE}
library(metafor)
library(devtools)
library(rotl)
library(ape)
library(orchaRd)
```

# Sensitivity analysis using leave-one-out

One way to assess the robustness of a meta-analytic model is by re-running the
model without one or more effect sizes and then comparing the results. A common
approach is to re-run the model multiple times, each time removing the effect
sizes from a different study. This allows you to see how the overall effect
estimate changes when each study is excluded.

This is a leave-one-out analysis (Nakagawa et al. 2023). In that example it 
leaves out one study at a time, but you can leave out any group you think can
have a large influence on the overall effect estimate.
In ecology and evolution meta-analyses, it can be useful to check how the model
changes when certain species are left out.

There are two functions for this: `leave_one_out()` and `orchard_leave1out()`.
The first one is to run the analysis, and the second one is to plot the results.


# Simple leave-one-out

Using the data from English and Uller (2016), here is how to run a leave-one-out
and plot the results. 

First, calculate the effect sizes and fit the model. 

```{r}
data(english)

# Create a new column with 'Author, year' format
english$author_year <- paste(english$Author, english$Year)


# We need to calculate the effect sizes, in this case d
english <- escalc(measure = "SMD",
                  n1i = NStartControl,
                  sd1i = SD_C,
                  m1i = MeanC,
                  n2i = NStartExpt,
                  sd2i = SD_E,
                  m2i = MeanE,
                  var.names = c("SMD", "vSMD"),
                  data = english)

english_MA <- rma.mv(yi = SMD,
                     V = vSMD,
                     random = list(~1 | StudyNo,
                                   ~1 | EffectID),
                    data = english)
```


`leave_one_out()` will re-run the model without one study at a time. 
The `group` argument is used to specify the grouping variable. In this case, 
we use the `author_year` variable:

```{r}
loo_test <- leave_one_out(english_MA, group = "author_year")
loo_test
```

We can plot the results with `orchard_leave1out()`.
This function takes the output of `leave_one_out()` and creates an orchard plot. 
The y-axis shows which study was left out. 

```{r, fig.height=10, fig.width=7}
orchard_leave1out(leave1out = loo_test,
                  xlab = "SMD",
                  ylab = "Study left out",
                  trunk.size = 1.2,
                  branch.size = 1.5,
                  alpha = 0.08,
                  legend.pos = "top.out")
```

- The dashed red lines show the 95% confidence interval for the overall effect size
in the original model. You can removed them by setting `ci_lines = FALSE`.
To change the color, use `ci_lines_color = "blue"`, 
or whatever color you want.

- The empty points, called 'ghost points', show the effect sizes left-out
of that model. You can removed them by setting `ghost_points = FALSE`.


The rest of the arguments are the same as in `orchard_plot()`.


# Leave-one-out in more complex models

There are cases in which the meta-analytic model is more complicated. 
`leave_one_out()` can handle some common cases that try to account for
non-independence:

- Variance-covariance matrix for sampling variances, 
implemented with `metafor::vcalc()`

- Phylogenetic matrix, calculated with `ape::vcv()`

- Robust variance estimation, implemented with `metafor::robust()`


To demonstrate how this works, we will use a subset of the Pottier data and
fit a fictitious model.



```{r, warning=FALSE, results='hide'}
data("pottier")
pottier_subset <- pottier[1:20, ]


# ----------------------------------------------------
# Get the phylo tree using `rotl`
unique_names <- unique(pottier_subset$search_string)
taxa <- rotl::tnrs_match_names(unique_names)

tree <- rotl::tol_induced_subtree(taxa$ott_id)
tree$tip.label <- rotl::strip_ott_ids(tree$tip.label)

# Compute the phylogenetic matrix using `ape` package
tree <- ape::compute.brlen(tree)
phylo_matrix <- ape::vcv(tree, corr = TRUE)

# ----------------------------------------------------
# Get the variance-covariance matrix
VCV <- metafor::vcalc(vi = Var_dARR,
                      cluster = study_ID,
                      obs = es_ID,
                      rho = 0.5,
                      data = pottier_subset)
```

```{r}
# Fictitious model just for the example
res <- rma.mv(yi = dARR,
              V = VCV,
              test = "t",
              random = list(~1 | study_ID,
                            ~1 | es_ID,
                            ~1 | species_ID,
                            ~1 | phylogeny),
              R = list(phylogeny = phylo_matrix),
              data = pottier_subset)

robust_res <- robust(res, cluster = study_ID)

mod_results(robust_res, group = "study_ID")
```

To include those correlations in the leave-one-out analysis, we need to pass
some extra arguments to the `leave_one_out()` function, so they can be re-calculated
in each iteration. To see the details of the arguments, run `?leave_one_out()`.

```{r}
# vcalc_args must be a list with the arguments for vcalc. They must be
# passed as strings, except for rho, which is a number.
vcalc_args <- list(vi = "Var_dARR",
                   cluster = "study_ID",
                   obs = "es_ID",
                   rho = 0.5)

# phylo_args must be a list with the arguments for vcv. They must be
# the phylogenetic tree object, and a string with the name of the 
# species names column that is used as random effect.
phylo_args <- list(tree = tree,
                   species_colname = "phylogeny")

# robust_args must be a list with the arguments for robust. For the moment,
# it only accepts cluster as a string, and clubSandwich as TRUE or FALSE.
robust_args <- list(cluster = "study_ID")

loo_test_phylo <- leave_one_out(res,
                                group = "phylogeny",
                                vcalc_args = vcalc_args,
                                phylo_args = phylo_args,
                                robust_args = robust_args)

loo_test_phylo
```

```{r}
orchard_leave1out(leave1out = loo_test_phylo,
                  xlab = "dARR",
                  ylab = "Species left out",
                  trunk.size = 1.2,
                  branch.size = 1.5,
                  alpha = 0.2,
                  legend.pos = "top.out")
```

Note that this time the leave-one-out was done by species, not by study.
You can leave out one study at a time and still use phylogeny:

```{r}
loo_test_stdy <- leave_one_out(res,
                          group = "study_ID",
                          vcalc_args = vcalc_args,
                          phylo_args = phylo_args,
                          robust_args = robust_args)
loo_test_stdy
```

Finally, you can use the output from leave_one_out() to plot the results
using ggplot2 if you like:

```{r}
library(ggplot2)

loo_test_phylo$mod_table |>
  ggplot(aes(x = name, y = estimate, ymin = lowerCL, ymax = upperCL)) +
  geom_pointrange() +
  coord_flip() 
```

